Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot Prediction?

### 2021-10-30 forecast/paper/paper.tex
Second push done
- Incorporate sentence update in intro from Jacob
- Take a careful pass through rest of the paper, various small
  edits or copyedits
- Restore some of the original discussion, but try to avoid having
  it sound like "use my method"

% Please add a significance statement to explain the relevance of your work
\significancestatement{Validated forecasting methodology should be a key 
  element in the public health response to any fast-moving epidemic or 
  pandemic. A widely-used model for predicting the future spread of a temporal
  process, based on past data, is an autoregressive (AR) model. While basic,
  such an AR model (properly trained) is already competitive with top models
  in operational use for COVID-19 forecasting. In this paper, we 
  exhibit five auxiliary indicators---based on de-identified medical insurance
  claims, self-reported symptoms via online surveys, and COVID-related Google
  searches---that further improve the predictive accuracy of an AR model in
  COVID-19 forecasting. The most substantial gains appear to be in quiescent  
  times; but the Google search indicator appears to also offer improvements
  during upswings in pandemic activity.   
}

@@

\begin{abstract}
  Short-term forecasts of traditional public health reporting streams (such as
  cases, hospitalizations, and deaths) are a key input to public health
  decision-making during a pandemic. Since early 2020, our research group has
  worked with data partners to collect, curate, and make publicly available
  numerous real-time COVID-19 indicators, providing multiple views of pandemic
  activity in the U.S.  This paper studies the utility of five such
  indicators---derived from de-identified medical insurance claims,
  self-reported symptoms from online surveys, and COVID-related Google search
  activity---from a forecasting perspective. For each indicator, we ask whether
  its inclusion in an autoregressive (AR) model leads to improved predictive
  accuracy relative to the same model excluding it. Such an AR model, without
  external features, is already competitive with many top COVID-19 forecasting
  models in use today. Our analysis reveals that (a) inclusion of each of these
  five indicators improves on the overall predictive accuracy of the AR model;
  (b) predictive gains are in general most pronounced during times in which
  COVID cases are trending in ``flat'' or ``down'' directions; (c) one
  indicator, based on Google searches, seems to be particularly helpful
  during ``up''  trends.      
\end{abstract}

@@

In the companion paper \cite{Reinhart:2021}, we analyze correlations between
various indicators and COVID case rates. These correlations are natural
summaries of the contemporaneous association between an indicator and COVID
cases, but they fall short of delivering a satisfactory answer to the question
that motivates the current article: is the information contained in an indicator
demonstrably useful for the prediction tasks we care about?
% For such a question,
% lagged correlations (e.g., measuring the correlation between an indicator and
% COVID case rates several days in the future) move us in the right direction, but
% still do not move us all the way there.
Demonstrating \textit{utility for
  prediction} is a much higher standard than simply asking about 
correlations; to be useful in forecast or hotspot models, an indicator must
provide relevant information that is not otherwise contained in past values of
the case rate series itself (see the pioneering work on Granger Causality
\cite{engle1987a,Granger:1969} as well as detailed discussion below). We
assess this directly by inspecting the difference in  
predictive performance of simple autoregressive models trained with and without
access to past values of a particular indicator.

@@

Careful handling of data revisions plays a key role in our analysis. Signals 
computed from surveillance streams are often subject to latency and/or 
revision. For example, a signal based on aggregated medical insurance claims may
be available after just a few days, but it can then be substantially revised
over the next several weeks as additional claims are submitted and/or processed
late. Correlations between such a signal and case rates calculated ``after the
fact'' (i.e., computed retrospectively, using the finalized values of this
signal) will not deliver an honest answer to the question of whether this signal
would have been useful in real time. Instead, we build predictive models using
only the data that would have been available \textit{as of} the prediction date,
and compare the ensuing predictions in terms of accuracy. The necessity of
real-time data for honest forecast evaluations has been recognized in
econometrics for a long time, see
\cite{howrey1978a, harvey1983a, mankiw1986a, trivellato1986a, diebold1991a,
  patterson1995a, koenig2003a, croushore2006a, clark2009a, faust2009a,
  croushore2011a}; but it appears to be far less acknowledged in public health,
despite its demonstrable importance.

@@

The indicators we focus on provide information not generally available from
standard public health reporting. Among the many auxiliary indicators collected
in the API, we study the following five:  
\begin{itemize}
\item Change Healthcare COVID-like illness (CHNG-CLI): The percentage of
  outpatient visits that are primarily about COVID-related symptoms, based on
  de-identified Change Healthcare claims data.
\item Change Healthcare COVID (CHNG-COVID): The percentage of outpatient visits
  with confirmed COVID-19, based on the same claims data.
\item COVID Trends and Impact Survey COVID-like illness in the community
  (CTIS-CLI-in-community): The estimated percentage of the population who know
  someone in their local community who is sick, based on Delphi's COVID Trends
  and Impact Survey, in partnership with Facebook.
\item Doctor Visits COVID-like illness (DV-CLI): The same as CHNG-CLI, but
  computed based on de-identified medical insurance claims from other health
  systems partners.  
\item Google search trends for anosmia and ageusia (Google-AA): A measure of  
  Google search volume for queries that relate to anosmia or ageusia (loss of
  smell or taste), based on Google's COVID-19 Search Trends data set.
\end{itemize}
In short, we choose these indicators because, conceptually speaking, they
measure aspects of an individual's disease progression that would plausibly
precede the occurence of (at worst, co-occur  with) the report of a positive
COVID-19 test, through standard public health reporting streams.

@@

As for geographic resolution, we consider the prediction of COVID-19 case rates
and hotspots aggregated at the level of an individual \textit{hospital referral
  region} (HRR). HRRs correspond to groups of counties in the United States
within the same hospital referral system. The Dartmouth Atlas of Healthcare 
Policy~\cite{DartmouthHRR}, defines these 306 regions based on a number of
characteristics. They are contiguous regions such that most of the hospital
services for the underlying population are performed by hospitals within the
region. Each HRR also contains at least one city where major procedures
(cardiovascular or neurological) are performed. The smallest HRR has a
population of about 125,000. While some are quite large (such as the one
containing Los Angeles, which has more than 10 million people), generally HRRs  
are much more homogenous in size than the (approximately) 3200 counties,  
and they serve as a nice middle ground in between counties and states.

@@

Lastly, our treatment of the Google-AA signal is different from the rest.
Because Google's team did not start publishing this signal until early
September, 2020, we do not have true vintage data before then.  Furthermore, the
latency of the signal was always at least one week through 2020.  However,
unlike the claims-based signals, there is no reason for revisions to occur after
initial publication, and furthermore, the latency of the signal is not an
unavoidable property of the data type, so we simply use finalized signal values,
with zero latency, in our analysis.

@@

Why do we define the response variables via 7-day averaging? The short answer 
is robustness: averaging stabilizes the case time series, and moderates
uninteresting artifacts like day-of-the-week effects in the series.  Note that
we can also equivalently view this (equivalent up to a constant factor) as
predicting the HRR-level case incidence rate \textit{summed} over some 7-day
period in the future, and predicting a binary variable derived from this.

@@

In our analysis, we let the prediction date $t$ run over each day in between
early/mid June and December 31, 2020.  The precise start date differs for 
forecasting and hotspots prediction; for each task it was chosen to be the
earliest date at which the data needed to train all models was available, which
ends up being (per our setup, with 21 days of training data and lagged values 
of signals for features, as we will detail shortly) June 9, 2020 for
forecasting, and June 16, 2020 for hotspot prediction. (The bottleneck here is
the CTIS-CLI-in-community signal, which does not exist before early April 2020,
when the survey was first launched). The end date was chosen again with a 
consideration to align both tasks as best as possible and to coincide with the
introduction of widely available vaccines.

@@

At prediction time, in order to avoid crossing violations (that is, for two
levels $\tau' > \tau$, the predicted quantile at level $\tau$ exceeds the
predicted quantile at level $\tau'$), we apply a simple post-hoc sorting.  See
Figure~\ref{fig:trajectory} for an example forecast.

@@

An important detail is that in hotspot prediction we remove all data from
training and evaluation where, on average, fewer than 30 cases (this refers to a
count, not a rate) are observed over the preceding 7 days. This avoids having to
make arbitrary calls for a hotspot (or lack thereof) based on small counts.

@@

For forecasting, we evaluate the probabilistic forecasts produced by the
quantile models in \eqref{eq:forecast_ar} and \eqref{eq:forecast_ar_x} using
\textit{weighted interval score} (WIS), a quantile-based scoring rule; see,
e.g., \cite{Gneiting:2007}.  WIS is a proper score, which means that its
expectation is minimized by the population quantiles of the target variable.
The use of WIS in COVID-19 forecast scoring is discussed in \cite{Bracher:2021};
WIS is also the main evaluation metric used in the COVID-19 Forecast Hub. More
broadly, the specific form of WIS used here is a standard metric in the
forecasting community for evaluating quantile-based probabilistic forecasts,
just as mean squared forecast error is standard for point forecasts.

WIS is typically defined for quantile-based forecasts where the quantile levels
are symmetric around 0.5.  This is the case for our choice in
\eqref{eq:quantile_levels}.  Let $F$ be a forecaster comprised of predicted
quantiles $q_\tau$ parametrized by a quantile level $\tau$.  In the case of
symmetric quantile levels, this is equivalent to a collection of central
prediction intervals $(\ell_\alpha, u_\alpha)$, parametrized by an exclusion
probability $\alpha$. The WIS of the forecaster $F$, evaluated at the target
variable $Y$, is defined by:
\begin{equation}
\label{eq:wis_intervals}
\WIS(F,Y) = \sum_\alpha \Big\{\alpha(u_\alpha - \ell_\alpha) + 2 \cdot
\dist(Y, [\ell_\alpha, u_\alpha])\Big\},  
\end{equation}
where $\dist(a, S)$ is the distance between a point $a$ and set $S$ (the
smallest distance between $a$ and an element of $S$).  Note that, corresponding  
to \eqref{eq:quantile_levels}, the exclusion probabilities are $\alpha \in
\{0.05,\ 0.2,\ 0.5,\ 1\}$, resulting in 4 terms in the above sum.  By 
straightforward algebra, it is not hard to see WIS has an alternative 
representation in terms of the predicted quantiles themselves:
\begin{equation}
\label{eq:wis_quantiles}
\WIS(F,Y) = 2 \sum_\tau \phi_\tau(Y - q_\tau), 
\end{equation}
where $\phi_\tau(x) = \tau |x|$ for $x \geq 0$ and $\phi_\tau(x) = (1-\tau)
|x|$ for $x<0$, which is often called the ``tilted absolute'' loss.  While
\eqref{eq:wis_quantiles} is more general (it can accomodate asymmetric quantile
levels), the first form in \eqref{eq:wis_intervals} is typically preferred in
presentation, as the score nicely decouples into a ``sharpness'' component
(first term in each summand) and an ``under/overprediction'' component (second
term in each summand).  But the second form given in \eqref{eq:wis_quantiles} is 
especially noteworthy in our current study because it reveals WIS to be the same
as the quantile regression loss that we use to train our forecasting models
(i.e., we fit by optimizing WIS averaged over the training data).

@@

Some signals have a significant amount of spatial heterogeneity, by which we
mean their values across different geographic locations are not comparable.
This is the case for the Google-AA signal (due to the way in which the
underlying search trends time series is self-normalized, see
\cite{GoogleSymptoms}) and the claims-based signals (due to market-share 
differences, and/or differences in health-seeking behavior).  Such spatial
heterogeneity likely hurts the performance of the predictive models that rely on
these signals, because we train the models on data pooled over all locations.
In the current paper, we do not attempt to address this issue (it is again a
topic of ongoing work in our group), and we simply note that location-specific
effects (or pre-processing to remove spatial bias) would likely improve the
performance of the models involving Google-AA and the claims-based indicators.

@@

Here, and in what follows, we will use ``AR'' to refer to the pure 
autoregressive model both in forecasting, \eqref{eq:forecast_ar}, and in hotspot
prediction, \eqref{eq:hotspot_ar} (the reference to the prediction task should
always be clear from the context). We will also use the name of an auxiliary 
indicator---namely ``CHNG-CLI'',  ``CHNG-COVID'', ``CTIS-CLI-in-community'',
``DV-CLI'', or ``Google-AA''---interchangeably with the model in forecasting, 
\eqref{eq:forecast_ar_x}, or hotspot prediction, \eqref{eq:hotspot_ar_x}, that
uses this particular indicator as a feature (the meaning should be clear from
the context).  So, for example, the CHNG-CLI model in forecasting is the one in  
\eqref{eq:forecast_ar_x} that sets $X_{\ell,t}$ to be the value of the CHNG-CLI
indicator at location $\ell$ and time $t$.  Finally, we use the term ``indicator
model'' to refer to any one of the ten models of the form
\eqref{eq:forecast_ar_x} or \eqref{eq:hotspot_ar_x} (five from each of the
forecasting and hotspot prediction tasks).

We begin with a summary of the high-level conclusions.

@@

\item In the same aggregate view, CHNG-COVID and DV-CLI offer the biggest 
  gains in both forecasting and hotspot prediction.  CHNG-CLI is
  inconsistent: it provides a big gain in hotspot prediction, but little gain 
  in forecasting (it seems to be hurt by a notable lack of robustness, due
  to backfill).  CTIS-CLI-in-community and Google-AA each provide decent 
  gains in forecasting and hotspot prediction.  The former's performance in 
  forecasting is notable in that it clearly improves on AR even at the largest
  ahead values.

\item In a more detailed analysis of forecasting performance, we find that the
  indicator models tend to be better than AR when case rates are flat or
  decreasing (most notable in CHNG-COVID and CTIS-CLI-in-community), but can be
  worse than AR when case rates are increasing (this is most notable in CHNG-CLI
  and DV-CLI). More rarely does an indicator model tend to beat AR when case
  rates are increasing, but there appears to be some evidence of this for the
  Google-AA model.  
  % In an increasing period, when an indicator model is worse than AR, it tends
  % to underpredict relative to AR (its median forecast is lower).  In a flat
  % or decreasing period, when an indicator model is better than AR, AR tends
  % to overpredict relative to it (the median AR forecast is higher).

@@

Code to reproduce all results (which uses the \texttt{evalcast} R package) can 
be found at
\url{https://github.com/cmu-delphi/covidcast-pnas/tree/main/forecast/code}.

@@

One might ask if the benefits we observe in forecasting and hotspot
prediction have anything to do with the actual indicator themselves. A plausible  
alternative explanation is that the indicators are simply providing \textit{implicit
  regularization} on top of the basic AR model, in the same way any noise
variable might, when we include them as lagged features in
\eqref{eq:forecast_ar_x} and \eqref{eq:hotspot_ar_x}.

To test this hypothesis, we reran all of the prediction experiments but with
$X_{\ell,t}$ in each indicator model replaced with suitable random noise
(bootstrap samples from a signal's history).  The results, shown and explained
more precisely in the Supplementary Materials, are vastly different (worse) than
the original set of results.  In both forecasting and hotspot prediction, the
``fake'' indicator models offered essentially no improvement over the pure AR
model, which---informally speaking---strongly rejects the implicit regularization
hypothesis.

On the topic of regularization, it is also worth noting that the use of 
$\ell_1$ regularization (tuned by cross-validation) in fitting any of the models
in \eqref{eq:forecast_ar}, \eqref{eq:forecast_ar_x}, \eqref{eq:hotspot_ar}, or 
\eqref{eq:hotspot_ar_x} did not generally improve their performance
(experiments not shown).  This is likely due to the fact that the number of
training samples is large compared to the number of features (6,426 training
samples and only 3--6 features). 
% DJM: and that proper CV requires longer training windows, exacerbating the
% nonstationarity, a scenario in which its behavior is poorly understood.

@@

Figure~\ref{fig:up_down_flat} shows histograms of the differences in WIS of the 
AR model and each indicator model, where we stratify these differences by
whether the target occurs during a period of increasing cases rates (up),
decreasing case rates (down), or flat case rates (flat). To define the
increasing period, we use the same definition we used for the hotspot task in
Table \ref{tab:analysis_tasks}.  Therefore all hotspots are ``up'', while all
non-hotspots are either ``flat'' or ``down''.  For the ``down'' scenario, we 
simply use the opposite of the hotspot definition: $Y_{\ell,t}$ decreases by
more than 20\% relative to the preceding week.

While the performance of all forecasters, including AR, will generally degrade
in up periods, different models exhibit different and interesting patterns.
CHNG-CLI, CHNG-COVID, Google-AA, and especially CTIS-CLI-in-community have
large right tails (displaying improvements over AR) during the down periods. 
Google-AA and CTIS-CLI-in-community have large right tails during the flat
periods. CHNG-CLI and DV-CLI have large left tails (poor forecasts relative to
AR) in flat and up periods.  Google-AA is the only model that outperforms
the AR model, on average, over up periods.  Overall, the indicators seem to help
more during flat or down periods than up periods, with the exception of
Google-AA.

@@

In the supplement, we define a quantitative score to measure the
leadingness of an indicator, at any time $t$ and any location $\ell$, based on
cross correlations to case rates over a short time window around $t$.
The higher this score, the greater it ``leads'' case activity.
%
This analysis is closely related to Granger Causality \cite{engle1987a} and
draws on a large body of prior work that measures leadingness in economic time-series 
\cite[e.g.,][]{yeats1972a,emerson1996a,koch1988a,sargent1977a,engle1987a,shiskin1968a,granger1969a,lahiri1991a,koopmans1947a,StockWatson1989,diebold1989a,mitchell1938a,auerbach1982a,hamilton1996a}.
%
Figure 
\ref{fig:leading} displays correlations between the leadingness score of an 
indicator and the WIS difference (AR model minus an indicator model), stratified
by whether the target is classified as up, down, or flat.  One would naturally
expect that the WIS difference would be positively correlated with leadingness.
Somewhat surprisingly, this relationship turns out to be strongest in down
periods and weakest in up periods.  In fact, it is very nearly the case that for
each indicator, the strength of correlations only decreases as we move from down
to flat to up  periods.  In the supplement, we extend this analysis by studying 
analogous ``laggingness'' scores, but we do not find as clear patterns.

@@

We reiterate the importance of using vintage data for rigorous backtesting. Data
sources that are relevant to public health surveillance are often subject to
revision, sometimes regularly (such as medical claims data) and sometimes
unpredictably (such as COVID-19 case reports).  When analyzing models that are
designed to predict future events, if we train these models and make predictions
using finalized data, then we are missing a big part of the story.  Not only
will our sense of accuracy be unrealistic, but certain models may degrade by a
greater or lesser extent when they are forced to work with vintage data, so 
backtesting them with finalized data may lead us to make modeling decisions that
are suboptimal for true test-time performance.

>>>>

% Please add a significance statement to explain the relevance of your work
\significancestatement{Validated forecasting methodology should be a vital  
  element in the public health response to any fast-moving epidemic or 
  pandemic. A widely-used model for predicting the future spread of a temporal
  process, based on past data, is an autoregressive (AR) model. While basic,
  such an AR model (properly trained) is already competitive with the top models
  in operational use for COVID-19 forecasting. In this paper, we exhibit five
  auxiliary indicators---based on de-identified medical insurance claims,
  self-reported symptoms via online surveys, and COVID-related Google
  searches---that further improve the predictive accuracy of an AR model in
  COVID-19 forecasting. The most substantial gains appear to be in quiescent   
  times; but the Google search indicator appears to also offer improvements
  during upswings in pandemic activity.   
}

@@

\begin{abstract}
  Short-term forecasts of traditional streams from public health reporting (such 
  as cases, hospitalizations, and deaths) are a key input to public health
  decision-making during a pandemic. Since early 2020, our research group has
  worked with data partners to collect, curate, and make publicly available
  numerous real-time COVID-19 indicators, providing multiple views of pandemic
  activity in the U.S.  This paper studies the utility of five such
  indicators---derived from de-identified medical insurance claims,
  self-reported symptoms from online surveys, and COVID-related Google search
  activity---from a forecasting perspective. For each indicator, we ask whether
  its inclusion in an autoregressive (AR) model leads to improved predictive
  accuracy relative to the same model excluding it. Such an AR model, without
  external features, is already competitive with many top COVID-19 forecasting
  models in use today. Our analysis reveals that (a) inclusion of each of these
  five indicators improves on the overall predictive accuracy of the AR model;
  (b) predictive gains are in general most pronounced during times in which
  COVID cases are trending in ``flat'' or ``down'' directions; (c) one
  indicator, based on Google searches, seems to be particularly helpful
  during ``up''  trends.      
\end{abstract}

@@

In the companion paper \cite{Reinhart:2021}, we analyze correlations between
various indicators and COVID case rates. These correlations are natural
summaries of the contemporaneous association between an indicator and COVID
cases, but they fall short of delivering a satisfactory answer to the question
that motivates the current article: is the information contained in an indicator
demonstrably useful for the prediction tasks we care about? Note that even
lagged correlations cannot deliver a complete answer. Demonstrating
\textit{utility for prediction} is a much higher standard than simply asking
about correlations; to be useful in forecast or hotspot models, an indicator
must provide relevant information that is not otherwise contained in past values
of the case rate series itself (cf.\ the pioneering work on Granger  causality
\cite{Granger:1969, engle1987a}, as well as the further references given 
below). We assess this directly by inspecting the difference in predictive  
performance of simple autoregressive models trained with and without access to  
past values of a particular indicator.

@@

Careful handling of data revisions plays a key role in our analysis. Signals 
computed from surveillance streams are often subject to latency and/or 
revision. For example, a signal based on aggregated medical insurance claims may
be available after just a few days, but it can then be substantially revised
over the next several weeks as additional claims are submitted and/or processed
late. Correlations between such a signal and case rates calculated ``after the
fact'' (i.e., computed retrospectively, using the finalized values of this
signal) will not deliver an honest answer to the question of whether this signal
would have been useful in real time. Instead, we build predictive models using
only the data that would have been available \textit{as of} the prediction date,
and compare the ensuing predictions in terms of accuracy. The necessity of
real-time data for honest forecast evaluations has been recognized in
econometrics for a long time, see \cite{howrey1978a, harvey1983a, mankiw1986a,
  trivellato1986a, diebold1991a, patterson1995a, koenig2003a, croushore2006a,
  clark2009a, faust2009a, croushore2011a}; but it is often overlooked in
epidemic forecasting despite its critical importance, see, e.g.,
\cite{chakraborty2018know}.

@@

The indicators we focus on provide information not generally available from
standard public health reporting. Among the many auxiliary indicators collected
in the API, we study the following five:  
\begin{itemize}
\item Change Healthcare COVID-like illness (CHNG-CLI): The percentage of
  outpatient visits that are primarily about COVID-related symptoms, based on
  de-identified Change Healthcare claims data.
\item Change Healthcare COVID (CHNG-COVID): The percentage of outpatient visits
  with confirmed COVID-19, based on the same claims data.
\item COVID Trends and Impact Survey COVID-like illness in the community
  (CTIS-CLI-in-community): The estimated percentage of the population who know
  someone in their local community who is sick, based on Delphi's COVID Trends
  and Impact Survey, in partnership with Facebook.
\item Doctor Visits COVID-like illness (DV-CLI): The same as CHNG-CLI, but
  computed based on de-identified medical insurance claims from other health
  systems partners.  
\item Google search trends for anosmia and ageusia (Google-AA): A measure of  
  Google search volume for queries that relate to anosmia or ageusia (loss of
  smell or taste), based on Google's COVID-19 Search Trends data set.
\end{itemize}
We choose these indicators because, conceptually speaking, they measure aspects
of an individual's disease progression that would plausibly precede the  
occurence of (at worst, co-occur  with) the report of a positive COVID-19 test,
through standard public health reporting streams.

@@

As for geographic resolution, we consider the prediction of COVID-19 case rates
and hotspots aggregated at the level of an individual \textit{hospital referral
  region} (HRR). HRRs correspond to groups of counties in the United States
within the same hospital referral system. The Dartmouth Atlas of Healthcare 
Policy~\cite{DartmouthHRR}, defines these 306 regions based on a number of
characteristics. They are contiguous regions such that most of the hospital
services for the underlying population are performed by hospitals within the
region. Each HRR also contains at least one city where major procedures
(cardiovascular or neurological) are performed. The smallest HRR has a
population of about 125,000. While some are quite large (such as the one
containing Los Angeles, which has more than 10 million people), generally HRRs  
are much more homogenous in size than the (approximately) 3200 U.S. counties,   
and they serve as a nice middle ground in between counties and states.

@@

Lastly, our treatment of the Google-AA signal is different from the rest.
Because Google's team did not start publishing this signal until early
September, 2020, we do not have true vintage data before then; the latency of 
the signal was always at least one week through 2020.  However, unlike the
claims-based signals, there is no reason for revisions to occur after initial
publication, and furthermore, the latency of the signal is not an unavoidable
property of the data type, so we simply use finalized signal values, with zero
latency, in our analysis.

@@

Why do we define the response variables via 7-day averaging? The short answer 
is robustness: averaging stabilizes the case time series, and moderates
uninteresting artifacts like day-of-the-week effects in the data.  We note
that we can also equivalently view this (equivalent up to a constant factor) as 
predicting the HRR-level case incidence rate \textit{summed} over some 7-day
period in the future, and predicting a binary variable derived from this.

@@

In our analysis, we let the prediction date $t$ run over each day in between
early/mid June and December 31, 2020.  The precise start date differs for 
forecasting and hotspots prediction; for each task it was chosen to be the
earliest date at which the data needed to train all models was available, which
ends up being (per our setup, with 21 days of training data and lagged values 
of signals for features, as we will detail shortly) June 9, 2020 for
forecasting, and June 16, 2020 for hotspot prediction. (The bottleneck here is
the CTIS-CLI-in-community signal, which does not exist before early April 2020,
when the survey was first launched).

@@

At prediction time, in order to avoid crossing violations (that is, for two
levels $\tau' > \tau$, the predicted quantile at level $\tau$ exceeds the
predicted quantile at level $\tau'$), we apply a simple post hoc sorting.  See 
Figure~\ref{fig:trajectory} for an example forecast.

@@

An important detail is that in hotspot prediction we remove all data from
training and evaluation where, on average, fewer than 30 cases (this refers to a
count, not a rate) are observed over the prior 7 days. This avoids having to
make arbitrary calls for a hotspot (or lack thereof) based on small counts.

@@

For forecasting, we evaluate the probabilistic forecasts produced by the
quantile models in \eqref{eq:forecast_ar} and \eqref{eq:forecast_ar_x} using
\textit{weighted interval score} (WIS), a quantile-based scoring rule; see,
e.g., \cite{Gneiting:2007}.  WIS is a proper score, which means that its
expectation is minimized by the population quantiles of the target variable.
The use of WIS in COVID-19 forecast scoring was proposed by \cite{Bracher:2021}; 
WIS is also the main evaluation metric used in the COVID-19 Forecast Hub. More
broadly, the specific form of WIS used here is a standard metric in the
forecasting community for evaluating quantile-based probabilistic forecasts,
just as mean squared forecast error is standard for point forecasts.

WIS is typically defined for quantile-based forecasts where the quantile levels
are symmetric around 0.5.  This is the case for our choice in
\eqref{eq:quantile_levels}.  Let $F$ be a forecaster comprised of predicted
quantiles $q_\tau$ parametrized by a quantile level $\tau$.  In the case of
symmetric quantile levels, this is equivalent to a collection of central
prediction intervals $(\ell_\alpha, u_\alpha)$, parametrized by an exclusion
probability $\alpha$. The WIS of the forecaster $F$, evaluated at the target
variable $Y$, is defined by:
\begin{equation}
\label{eq:wis_intervals}
\WIS(F,Y) = \sum_\alpha \Big\{\alpha(u_\alpha - \ell_\alpha) + 2 \cdot
\dist(Y, [\ell_\alpha, u_\alpha])\Big\},  
\end{equation}
where $\dist(a, S)$ is the distance between a point $a$ and set $S$ (the
smallest distance between $a$ and an element of $S$).  Note that, corresponding  
to \eqref{eq:quantile_levels}, the exclusion probabilities are $\alpha \in
\{0.05,\ 0.2,\ 0.5,\ 1\}$, resulting in 4 terms in the above sum.  By 
straightforward algebra, it is not hard to see WIS has an alternative 
representation in terms of the predicted quantiles themselves:
\begin{equation}
\label{eq:wis_quantiles}
\WIS(F,Y) = 2 \sum_\tau \phi_\tau(Y - q_\tau), 
\end{equation}
where $\phi_\tau(x) = \tau |x|$ for $x \geq 0$ and $\phi_\tau(x) = (1-\tau)
|x|$ for $x<0$, which is often called the ``tilted absolute'' loss.  While
\eqref{eq:wis_quantiles} is more general (it can accomodate asymmetric quantile
levels), the first form in \eqref{eq:wis_intervals} is typically preferred in
presentation, as the score nicely decouples into a ``sharpness'' component
(first term in each summand) and an ``under/overprediction'' component (second
term in each summand).  But the second form given in \eqref{eq:wis_quantiles} is 
especially noteworthy in our current study because it reveals WIS to be the same
as the quantile regression loss that we use to train our forecasting models.
%(i.e., we fit by optimizing WIS averaged over the training data).

@@

Some signals have a significant amount of spatial heterogeneity, by which we
mean their values across different geographic locations are not comparable.
This is the case for the Google-AA signal (due to the way in which the
Google search trends time series is self-normalized, see
\cite{GoogleSymptoms}) and the claims-based signals (due to market-share 
differences, and/or differences in health-seeking behavior).  Such spatial
heterogeneity likely hurts the performance of the predictive models that rely on
these signals, because we train the models on data pooled over all locations.
In the current paper, we do not attempt to address this issue, and we simply
note that location-specific effects (or pre-processing to remove spatial bias)
would likely improve the performance of the models involving Google-AA and the
claims-based indicators.

@@

Here, and in what follows, we will use ``AR'' to refer to the pure 
autoregressive model both in forecasting, \eqref{eq:forecast_ar}, and in hotspot
prediction, \eqref{eq:hotspot_ar} (the reference to the prediction task should
always be clear from the context). We will also use the name of an auxiliary 
indicator---namely ``CHNG-CLI'',  ``CHNG-COVID'', ``CTIS-CLI-in-community'',
``DV-CLI'', or ``Google-AA''---interchangeably with the model in forecasting, 
\eqref{eq:forecast_ar_x}, or hotspot prediction, \eqref{eq:hotspot_ar_x}, that
uses this same indicator as a feature (the meaning should be clear from
the context).  So, for example, the CHNG-CLI model in forecasting is the one in  
\eqref{eq:forecast_ar_x} that sets $X_{\ell,t}$ to be the value of the CHNG-CLI
indicator at location $\ell$ and time $t$.  Finally, we use the term ``indicator
model'' to refer to any one of the ten models of the form
\eqref{eq:forecast_ar_x} or \eqref{eq:hotspot_ar_x} (five from each of the
forecasting and hotspot prediction tasks).

Below is summary of the high-level conclusions.

@@

\item In the same aggregate view, CHNG-COVID and DV-CLI offer the biggest 
  gains in both forecasting and hotspot prediction.  CHNG-CLI is
  inconsistent: it provides a large gain in hotspot prediction, but little gain  
  in forecasting (it seems to be hurt by a notable lack of robustness, due
  to backfill).  CTIS-CLI-in-community and Google-AA each provide decent 
  gains in forecasting and hotspot prediction.  The former's performance in 
  forecasting is notable in that it clearly improves on AR even at the largest
  ahead values.

\item In a more detailed analysis of forecasting performance, we find that the
  indicator models tend to be better than AR when case rates are flat or
  decreasing (most notable in CHNG-COVID and CTIS-CLI-in-community), but can be
  worse than AR when case rates are increasing (this is most notable in CHNG-CLI
  and DV-CLI). More rarely does an indicator model tend to beat AR when case
  rates are increasing, but there appears to be evidence of this for the
  Google-AA model.   
  % In an increasing period, when an indicator model is worse than AR, it tends
  % to underpredict relative to AR (its median forecast is lower).  In a flat
  % or decreasing period, when an indicator model is better than AR, AR tends
  % to overpredict relative to it (the median AR forecast is higher).

@@

Code to reproduce all results can be found at
\url{https://github.com/cmu-delphi/covidcast-pnas/tree/main/forecast/code}.

@@

One might ask if the benefits observed in forecasting and hotspot
prediction have anything to do with the actual auxiliary indicators
themselves. A plausible alternative explanation is that the indicators are
just providing \textit{implicit regularization} on top of the basic AR model,
in the same way any noise variable might, if we were to use it to create lagged
features in \eqref{eq:forecast_ar_x} and \eqref{eq:hotspot_ar_x}.

To test this hypothesis, we reran all of the prediction experiments but with
$X_{\ell,t}$ in each indicator model replaced by suitable random noise
(bootstrap samples from a signal's history).  The results, shown and explained
more precisely in the Supplement, are vastly different (i.e., worse) than the
original set of results.  In both forecasting and hotspot prediction, the
``fake'' indicator models offer essentially no improvement over the pure AR
model, which---informally speaking---strongly rejects the implicit
regularization hypothesis.

On the topic of regularization, it is also worth noting that the use of 
$\ell_1$ regularization (tuned using cross-validation) in fitting any of the
models in \eqref{eq:forecast_ar}, \eqref{eq:forecast_ar_x},
\eqref{eq:hotspot_ar}, or  \eqref{eq:hotspot_ar_x} did not generally improve
their performance (experiments not shown).  This is likely due to the fact that
the number of training samples is large compared to the number of features
(6,426 training samples and only 3--6 features).  
% DJM: and that proper CV requires longer training windows, exacerbating the
% nonstationarity, a scenario in which its behavior is poorly understood.

@@

Figure~\ref{fig:up_down_flat} shows histograms of the differences in WIS of the 
AR model and each indicator model, where we stratify these differences by
whether the target occurs during a period of increasing cases rates (up),
decreasing case rates (down), or flat case rates (flat). To define the
increasing period, we use the same definition we used for the hotspot task in
Table \ref{tab:analysis_tasks}.  Therefore, all hotspots are labeled ``up'',
while all non-hotspots are either ``flat'' or ``down''.  For the ``down''
scenario, we simply use the opposite of the hotspot definition: $Y_{\ell,t}$
decreases by more than 20\% relative to the preceding week.

While the performance of all models, including AR, generally degrades in up
periods, different models exhibit different and interesting patterns. CHNG-CLI,
CHNG-COVID, Google-AA, and especially CTIS-CLI-in-community have large right
tails (showing improvements over AR) during the down periods.  Google-AA and
CTIS-CLI-in-community have large right tails during the flat periods. CHNG-CLI
and DV-CLI have large left tails (poor forecasts relative to AR) in flat and up
periods.  Google-AA is the only model that outperforms the AR model, on average,
in up periods.  Overall, the indicators seem to help more during flat or down
periods than during up periods, with the exception of Google-AA.

@@

In the supplement, we define a quantitative score to measure the 
leadingness of an indicator, at any time $t$ and any location $\ell$, based on 
cross correlations to case rates over a short time window around $t$.
The higher this score, the greater it ``leads'' case activity. This analysis is
closely related to Granger causality \cite{engle1987a} and draws on a large
body of prior work that measures leadingness in economic time-series, e.g., 
\cite{mitchell1938a, koopmans1947a, shiskin1968a, granger1969a, yeats1972a, 
  sargent1977a, auerbach1982a, koch1988a, StockWatson1989, diebold1989a,
  lahiri1991a, emerson1996a, hamilton1996a}. Figure \ref{fig:leading} displays
correlations between the leadingness score of an indicator and the WIS
difference (AR model minus an indicator model), stratified by whether the target
is classified as up, down, or flat.  One would naturally expect that the WIS
difference would be positively correlated with leadingness.  Somewhat
surprisingly, this relationship turns out to be strongest in down periods and
weakest in up periods.  In fact, it is very nearly the case that for each
indicator, the strength of correlations only decreases as we move from down to
flat to up periods.  In the supplement, we extend this analysis by studying
analogous ``laggingness'' scores, but we do not find as clear patterns.

@@

We reiterate the importance of using vintage data for rigorous backtesting. Data
sources that are relevant to public health surveillance are often subject to
revision, sometimes regularly (as in medical claims data) and sometimes
unpredictably (such as COVID-19 case reports).  When analyzing models that are
designed to predict future events, if we train these models and make predictions
using finalized data, then we are missing a big part of the story.  Not only
will our sense of accuracy be unrealistic, but certain models may degrade by a
greater or lesser extent when they are forced to work with vintage data, so 
backtesting with finalized data may lead us to make modeling decisions that are
suboptimal for true test-time performance.
